---
title: "Dogs vs. Cats classification with mxnet and R"
author: "Andrey Ogurtsov (https://github.com/statist-bhfz/)"
date: "February 25, 2017"
---

## 1. Packages and prerequisites

Ubuntu 16, **mxnet** 0.9.4 (compiled with GPU support), **imager** for image processind, **abind** for manipulations with arrays. It is almost end-to-end R solution for Kaggle competition https://www.kaggle.com/c/dogs-vs-cats-redux-kernels-edition/, we will use Python only for creating .rec-files.

Thanks to [jeremiedb](https://github.com/jeremiedb), my code for fine-tuning is largely based on his [answers](https://github.com/dmlc/mxnet/issues/4817).

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

```{r}
library(imager)
library(mxnet)
library(abind)
```


## 2. Image processing

### 2.1. Renaming train files

```{r}
files <- list.files("train")
old_names <- sapply(files, strsplit, split = ".", fixed = TRUE)
max_length <- max(sapply(old_names, function(x) nchar(x[[2]])))
zeros <- max_length - sapply(old_names, function(x) nchar(x[[2]]))
zeros <- sapply(zeros, function(x) paste(rep(0, x), collapse = ""))
new_names <- Map(function(x, y) {paste0("./train/", 
                                        x[1], 
                                        "/",
                                        y,
                                        x[2],
                                        ".jpg")},
                 x = old_names, y = zeros
                 )

# Full names
files <- paste0("./train/", files)

dir.create("./train/cat")
dir.create("./train/dog")

# New names will be in 00001.jpg format
Map(function(x, y) file.rename(from = x, to = y), files, new_names)
```

### 2.2. Train images: 224x224, padded with empty space

```{r}
files <- list.files("train", recursive = TRUE)
new_names <- paste0("train_pad_224x224/", files)
files <- paste0("./train/", files)
dir.create("./train_pad_224x224/")
dir.create("./train_pad_224x224/cat")
dir.create("./train_pad_224x224/dog")

padImage <- function(x) {
    long_side <- max(dim(x)[1:2])
    short_side <- min(dim(x)[1:2])
    pad_img <- pad(x, 
                   nPix = long_side - short_side,
                   axes = ifelse(dim(x)[1] < dim(x)[2], "x", "y"))
    return(pad_img)
}

Map(function(x, y) {
    pad_img <- padImage(load.image(x))
    res_img <- resize(pad_img,  size_x = 224, size_y = 224)
    imager::save.image(res_img, y)
}, x = files, y = new_names)
```

### 2.3. Renaming test files

```{r}
files <- list.files("test")
max_length <- max(sapply(files, nchar))
zeros <- max_length - sapply(files, nchar)
zeros <- sapply(zeros, function(x) paste(rep(0, x), collapse = ""))
newnames <- paste0("./test/", zeros, files)

files <- paste0("./test/", files)

Map(function(x, y) file.rename(from = x, to = y), files, newnames)
```


### 2.4. Test images: 224x224, padded with empty space

```{r}
files <- list.files("test")
new_names <- paste0("test_pad_224x224/", files)
files <- paste0("./test/", files)
dir.create("./test_pad_224x224/")

Map(function(x, y) {
    pad_img <- padImage(load.image(x))
    res_img <- resize(pad_img,  size_x = 224, size_y = 224)
    imager::save.image(res_img, y)
}, x = files, y = new_names)
```

### 2.5. Creating .rec files

```{bash, eval = FALSE}
python ~/mxnet/tools/im2rec.py --list=1 --recursive=1 --train-ratio=0.8 cats_dogs train_pad_224x224
python ~/mxnet/tools/im2rec.py --num-thread=4 --pass-through=1 cats_dogs_train.lst train_pad_224x224
python ~/mxnet/tools/im2rec.py --num-thread=4 --pass-through=1 cats_dogs_val.lst train_pad_224x224
```


## 3. Iterators

```{r}
get_iterator <- function(data_shape, 
						 train_data, 
						 val_data, 
						 batch_size = 128) {
    train <- mx.io.ImageRecordIter(
    	path.imgrec = train_data,
    	batch.size  = batch_size,
        data.shape  = data_shape,
    	rand.crop   = TRUE,
    	rand.mirror = TRUE)
  
    val <- mx.io.ImageRecordIter(
    	path.imgrec = val_data,
    	batch.size  = batch_size,
    	data.shape  = data_shape,
    	rand.crop   = FALSE,
    	rand.mirror = FALSE
    	)
 
  return(list(train = train, val = val))
}
```


```{r}
data  <- get_iterator(data_shape = c(224, 224, 3),
         train_data = "/media/andrey/Data/KAGGLE/cats_dogs/cats_dogs_train.rec",
         val_data   = "/media/andrey/Data/KAGGLE/cats_dogs/cats_dogs_val.rec",
         batch_size = 8)
train <- data$train
val   <- data$val
```


## 4. Load pretrained model

Model from http://data.dmlc.ml/models/imagenet/
Last fully connected layes for 1000 classes replaced with new layer for 2 classes.


```{r}
inception_bn <- mx.model.load("models/inception_bn/Inception-BN", 
							  iteration = 126)

symbol <- inception_bn$symbol
# check symbol$arguments for layer names
internals <- symbol$get.internals()
outputs <- internals$outputs

flatten <- internals$get.output(which(outputs == "flatten_output"))

new_fc <- mx.symbol.FullyConnected(data = flatten, 
                                   num_hidden = 2, 
                                   name = "fc1") 
                        # set name to original name in symbol$arguments
new_soft <- mx.symbol.SoftmaxOutput(data = new_fc, 
                                    name = "softmax")
                        # set name to original name in symbol$arguments

arg_params_new <- mxnet:::mx.model.init.params(
	symbol = new_soft, 
    input.shape = c(224, 224, 3, 8), 
    initializer = mxnet:::mx.init.uniform(0.1), 
    ctx = mx.gpu(0)
	)$arg.params
fc1_weights_new <- arg_params_new[["fc1_weight"]]
fc1_bias_new <- arg_params_new[["fc1_bias"]]

arg_params_new <- inception_bn$arg.params

arg_params_new[["fc1_weight"]] <- fc1_weights_new 
arg_params_new[["fc1_bias"]] <- fc1_bias_new 
```


## 5. Train (fine-tune) model

```{r}
model <- mx.model.FeedForward.create(
  symbol             = new_soft,
  X                  = train,
  eval.data          = val,
  ctx                = mx.gpu(0),
  eval.metric        = mx.metric.accuracy,
  num.round          = 1,
  learning.rate      = 0.05,
  momentum           = 0.9,
  wd                 = 0.00001,
  kvstore            = "local",
  array.batch.size   = 128,
  epoch.end.callback = mx.callback.save.checkpoint("inception_bn"),
  batch.end.callback = mx.callback.log.train.metric(150),
  initializer        = mx.init.Xavier(factor_type = "in", magnitude = 2.34),
  optimizer          = "sgd",
  arg.params         = arg_params_new,
  aux.params         = inception_bn$aux.params
)
```

```{r}
model <- mx.model.load("inception_bn", 1)
```

Continue training with decreased speed (`learning.rate = 0.03`):

```{r}
model <- mx.model.FeedForward.create(
  symbol             = model$symbol,
  X                  = train,
  eval.data          = val,
  ctx                = mx.gpu(0),
  eval.metric        = mx.metric.accuracy,
  num.round          = 5,
  learning.rate      = 0.03,
  momentum           = 0.9,
  wd                 = 0.00001,
  kvstore            = "local",
  array.batch.size   = 100,
  epoch.end.callback = mx.callback.save.checkpoint("inception_bn"),
  batch.end.callback = mx.callback.log.train.metric(150),
  initializer        = mx.init.Xavier(factor_type = "in", magnitude = 2.34),
  optimizer          = "sgd",
  arg.params         = model$arg.params, 
  aux.params         = model$aux.params
)
```

```{r}
model <- mx.model.load("inception_bn", 1)
```

My R session crashed after each iteration, so I made some iterations manually.


## 6. Make predictions

```{r}
preprocImage<- function(src,              # URL or file location
						height = 224,        
						width = 224,  
						num_channels = 3, # 3 for RGB, 1 for grayscale
						mult_by = 1,      # set to 255 for normalized image
						crop = FALSE) {   # no crop by default
	
	im <- load.image(src)
	
	if (crop) {
		shape <- dim(im)
        short_edge <- min(shape[1:2])
		xx <- floor((shape[1] - short_edge) / 2)
        yy <- floor((shape[2] - short_edge) / 2) 
        im <- crop.borders(im, xx, yy)
	}
	
	resized <- resize(im,  size_x = width, size_y = height)
	arr <- as.array(resized) * mult_by
	dim(arr) <- c(width, height, num_channels, 1)
	return(arr)
} 
```

```{r}
files <- list.files("test_pad_224x224/")
files <- paste0("./test_pad_224x224/", files)

# ind <- seq(1, 12500, 1250) 
# probs <- numeric()
# for (i in ind) {
#    images <- lapply(files[i:i+1249], preprocImage, mult_by = 255)
#    images <- do.call(abind, images)
#    probs[i:i+1249] <- predict(model, X = images, ctx = mx.gpu(0))
# }

files <- split(files, rep(1:1250, each = 10))
probs <- lapply(files, function(x) {
    images <- lapply(x, preprocImage, mult_by = 255)
    images <- do.call(abind, images)
    probs <- predict(model, X = images, ctx = mx.gpu(0))
})
saveRDS(probs, "probs.rds")
probs <- t(do.call(cbind, probs))

preds <- data.frame(id = 1:12500, label = probs[, 2])
write.csv(preds, "subm.csv", row.names = FALSE, quote = FALSE)
```
