Char RNN Example
=============================================

This example aims to show how to use lstm model to build a char level language model, and generate text from it. We use a tiny shakespeare text for demo purpose.

Data can be found at https://github.com/dmlc/web-data/tree/master/mxnet/tinyshakespeare. 

Preface
-------
This tutorial is written in Rmarkdown.
- You can directly view the hosted version of the tutorial from [MXNet R Document](http://mxnet.readthedocs.org/en/latest/package/r/CharRnnModel.html)
- You can find the download the Rmarkdown source from [here](https://github.com/dmlc/mxnet/blob/master/R-package/vignettes/CharRnnModel.Rmd)

Load Data 
---------
First of all, load in the data and preprocess it.
```{r}
require(mxnet)
```
Set basic network parameters.
```{r}
batch.size = 32
seq.len = 32
num.hidden = 256
num.embed = 256
num.lstm.layer = 2
num.round = 3
learning.rate= 0.1
wd=0.00001
clip_gradient=1
update.period = 1
```
download the data.
```{r}
download.data <- function(data_dir) {
    dir.create(data_dir, showWarnings = FALSE)
    if (!file.exists(paste0(data_dir,'input.txt'))) {
        download.file(url='https://raw.githubusercontent.com/dmlc/web-data/master/mxnet/tinyshakespeare/input.txt',
                      destfile=paste0(data_dir,'input.txt'), method='wget')
    }
}
```
Make dictionary from text.
```{r}
make.dict <- function(text, max.vocab=10000) {
    text <- strsplit(text, '')
    dic <- list()
    idx <- 1
    for (c in text[[1]]) {
        if (!(c %in% names(dic))) {
            dic[[c]] <- idx
            idx <- idx + 1
        }
    }
    if (length(dic) == max.vocab - 1)
        dic[["UNKNOWN"]] <- idx
    cat(paste0("Total unique char: ", length(dic), "\n"))
    return (dic)
}
```
Transfer text into data feature.
```{r}
make.data <- function(file.path, seq.len=32, max.vocab=10000, dic=NULL) {
    fi <- file(file.path, "r")
    text <- paste(readLines(fi), collapse="\n")
    close(fi)

    if (is.null(dic))
        dic <- make.dict(text, max.vocab)
    lookup.table <- list()
    for (c in names(dic)) {
        idx <- dic[[c]]
        lookup.table[[idx]] <- c 
    }

    char.lst <- strsplit(text, '')[[1]]
    num.seq <- as.integer(length(char.lst) / seq.len)
    char.lst <- char.lst[1:(num.seq * seq.len)]
    data <- array(0, dim=c(seq.len, num.seq))
    idx <- 1
    for (i in 1:num.seq) {
        for (j in 1:seq.len) {
            if (char.lst[idx] %in% names(dic))
                data[j, i] <- dic[[ char.lst[idx] ]]-1
            else {
                data[j, i] <- dic[["UNKNOWN"]]-1
            }
            idx <- idx + 1
        }
    }
    return (list(data=data, dic=dic, lookup.table=lookup.table))
}
```
Move tail text.
```{r}
drop.tail <- function(X, batch.size) {
    shape <- dim(X)
    nstep <- as.integer(shape[2] / batch.size)
    return (X[, 1:(nstep * batch.size)])
}
```
get the label of X
```{r}
get.label <- function(X) {
    label <- array(0, dim=dim(X))
    d <- dim(X)[1]
    w <- dim(X)[2]
    for (i in 0:(w-1)) {
        for (j in 1:d) {
            label[i*d+j] <- X[(i*d+j)%%(w*d)+1]
        }
    }
    return (label)
}
```
get training data and eval data
```{r}
download.data("./data/")
ret <- make.data("./data/input.txt", seq.len=seq.len)
X <- ret$data
dic <- ret$dic
lookup.table <- ret$lookup.table

vocab <- length(dic)

shape <- dim(X)
train.val.fraction <- 0.9
size <- shape[2]

X.train.data <- X[, 1:as.integer(size * train.val.fraction)]
X.val.data <- X[, -(1:as.integer(size * train.val.fraction))]
X.train.data <- drop.tail(X.train.data, batch.size)
X.val.data <- drop.tail(X.val.data, batch.size)

X.train.label <- get.label(X.train.data)
X.val.label <- get.label(X.val.data)

X.train <- list(data=X.train.data, label=X.train.label)
X.val <- list(data=X.val.data, label=X.val.label)
```

Training Model
--------------
In `mxnet`, we have a function called `mx.lstm` so that users can build a general lstm model. 

```{r}
model <- mx.lstm(X.train, X.val, 
                 ctx=mx.cpu(),
                 num.round=num.round, 
                 update.period=update.period,
                 num.lstm.layer=num.lstm.layer, 
                 seq.len=seq.len,
                 num.hidden=num.hidden, 
                 num.embed=num.embed, 
                 num.label=vocab,
                 batch.size=batch.size, 
                 input.size=vocab,
                 initializer=mx.init.uniform(0.1), 
                 learning.rate=learning.rate,
                 wd=wd,
                 clip_gradient=clip_gradient)

```
Setting the parameters ctx=mx.gpu(0) and num.round=5 can get the following result.
```
Epoch [31] Train: NLL=3.47213018872144, Perp=32.2052727363657
...
Epoch [961] Train: NLL=2.32060007657895, Perp=10.181782322355
Iter [1] Train: Time: 186.397065639496 sec, NLL=2.31135356537961, Perp=10.0880702804858
Iter [1] Val: NLL=1.94184484060012, Perp=6.97160060607419
Epoch [992] Train: NLL=1.84784553299322, Perp=6.34613225095329
...
Epoch [1953] Train: NLL=1.70175791172558, Perp=5.48357857093351
Iter [2] Train: Time: 188.929051160812 sec, NLL=1.70103940328978, Perp=5.47963998859367
Iter [2] Val: NLL=1.74979316010449, Perp=5.75341251767988
...
Epoch [2914] Train: NLL=1.54738185300295, Perp=4.69915099483974
Iter [3] Train: Time: 185.425321578979 sec, NLL=1.54604189517013, Perp=4.69285854740519
Iter [3] Val: NLL=1.67780240235925, Perp=5.35377758479576
Epoch [2945] Train: NLL=1.48868466087876, Perp=4.43126307034767
...
Iter [4] Train: Time: 185.487086296082 sec, NLL=1.4744973925858, Perp=4.36883940994296
Iter [4] Val: NLL=1.64488167325603, Perp=5.18039689118454
Epoch [3937] Train: NLL=1.46355541021581, Perp=4.32129622881604
...
Epoch [4898] Train: NLL=1.42900458455642, Perp=4.17454171976281
Iter [5] Train: Time: 185.070136785507 sec, NLL=1.42909226256273, Perp=4.17490775130428
Iter [5] Val: NLL=1.62716655804022, Perp=5.08943365437187

```
Inference from model
--------------------
helper function for random sample.
```{r}
cdf <- function(weights) {
    total <- sum(weights)
    result <- c()
    cumsum <- 0
    for (w in weights) {
        cumsum <- cumsum+w
        result <- c(result, cumsum / total)
    }
    return (result)
}

search.val <- function(cdf, x) {
    l <- 1
    r <- length(cdf) 
    while (l <= r) {
        m <- as.integer((l+r)/2)
        if (cdf[m] < x) {
            l <- m+1
        } else {
            r <- m-1
        }
    }
    return (l)
}
choice <- function(weights) {
    cdf.vals <- cdf(as.array(weights))
    x <- runif(1)
    idx <- search.val(cdf.vals, x)
    return (idx)
}
```
we can use random output or fixed output by choosing largest probability.
```{r}
make.output <- function(prob, sample=FALSE, temperature=1.) {
    if (!sample) {
        idx <- which.max(as.array(prob))
    }
    else {
        scale_prob <- mx.nd.clip(prob, 1e-6, 1 - 1e-6)
        rescale <- mx.nd.exp(mx.nd.log(scale_prob) / temperature)
        rescale <- rescale / (as.array(mx.nd.sum(rescale))[1])
        idx <- choice(rescale)
    }
    return (idx)

}
```

In `mxnet`, we have a function called `mx.lstm.inference` so that users can build a inference from lstm model and then use function `mx.lstm.forward` to get forward output from the inference.
Build inference from model.
```{r}
infer.model <- mx.lstm.inference(num.lstm.layer=num.lstm.layer,
                                 input.size=vocab,
                                 num.hidden=num.hidden,
                                 num.embed=num.embed,
                                 num.label=vocab,
                                 arg.params=model$arg.params,
                                 ctx=mx.cpu())
```
generate a sequence of 75 chars using function `mx.lstm.forward`.
```
start <- 'a'
seq.len <- 75
random.sample <- TRUE

last.id <- dic[[start]]
out <- "a"
for (i in (1:(seq.len-1))) {
    input <- c(last.id-1)
    ret <- mx.lstm.forward(infer.model, input, FALSE)
    infer.model <- ret$model
    prob <- ret$prob
    last.id <- make.output(prob, random.sample)
    out <- paste0(out, lookup.table[[last.id]])
}
cat (paste0(out, "\n"))
```
The result:
```
ah not a drobl greens
Settled asing lately sistering sounted to their hight
```